---
title: "ENGSCI 314 Assignment 3"
author: "Your Name and ID number here"
date: 'Due Date: 1pm Tuesday, 29th May 2018'
output:
  html_document:
    fig_caption: yes
  pdf_document: default
  word_document: default
---
```{r echo=FALSE}
## Do not delete this!
## It loads the s20x library for you. If you delete it 
## your document may not compile it.
require(s20x)
```

##**Question 1**

###Question of interest/goal of the study






### Read in and inspect the data:
```{r,fig.height=6,fig.width=6}
game.df=read.table("game14.txt",header=TRUE)
pairs20x(game.df[1:4])

```





### Fit model and check assumptions
```{r,fig.height=3.8,fig.width=6}





```

### Create a plot with 'best' model superimposed on top of data.
```{r,fig.height=5,fig.width=6}





```

### Rotate model for final comparison
```{r,fig.height=3.8,fig.width=6}



```


### Method and Assumption Checks






### Executive Summary






****

##**Question 2**

### Question of interest/goal of the study





### Read in and inspect the data:
```{r,fig.height=5,fig.width=6}
iron.df=read.table("Iron.txt",header=TRUE)
interactionPlots(Iron~Pot+Food,data=iron.df)

```







### Fit model and check assumptions
```{r,fig.height=3.5,fig.width=6}




```

### Method and Assumption Checks








### Executive Summary








